Analysis

### 2019-09-09 reports/manuscript.Rmd
minor text changes

author: 
  - name          : "Filip Děchtěrenko"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Hybernská 8, 11000, Prague 1"
    email         : "filip.dechterenko@gmail.com"
  - name          : "Jiří Lukavský"
    affiliation   : "2"

@@

authornote: |
  Filip Děchtěrenko, Department of Psychology, Faculty of Arts, Charles University, Celetná 20, 110 00, Prague, Czech Republic. Institute of Psychology, Czech Academy of Sciences, Hybernská 8, 110 00, Prague, Czech Republic. Jiří Lukavský, Institute of Psychology, Czech Academy of Sciences, Hybernská 8, 110 00, Prague, Czech Republic.

@@

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_docx
---

@@

World around us is full of information to process and store. When studying visual perception, it has been observed that visual memory has massive capacity for complex stimuli [@standing1970, @standing1973]. Recent studies showed that we are able to store not only shallow representation of the objects, but they are able to store many details [@brady2008] and similar findings were observed for scenes [@konkle2010b]. 
Although the capacity is large, it is an open question, how is it organised. The high recollection accuracy get lower when using yes/no task instead of 2AFC [@Cunningham2015]. Moreover, recent results show that the memory is much more limited in detail than it seems [@Draschkow2019].

How is the long-term memory organized? One possible explanation would be to create a coding scheme, using which the memories will be organised [@brady2011], similarly as in working memory [@brady2009]. This coding scheme is created during our lifetime with the formation of semantic categories [@Kemp2009]. In the case of objects, the possible candidate is a hierarchical structure from simple features through larger meaningful groups to whole objects [@brady2001, @Ullman2007]. In the case of scenes is the situation more complicated as not only the objects in the scene should be represented, but their spatial arrangement as well as the gist of the scene should be stored [@Oliva2005]. The organisation of whole scenes in the memory remains an open question.

One possible approach to this problem could be through formation of categories. The semantic categories in our mind are formed to match the categories of outside world and to be economic with terms of used properties to describe the structure [@rosch1978]. The category is formed around a prototype - a typical exemplar, which shows the maximal resemblance with rest of the members in the category, while having a low resemblance with other categories [@rosch1975b]. The prototypes are created not only for semantic categories, but also for visual stimuli. There are two main theories around visual prototype formation. The prototype-distance model [@Posner1970] states that the prototype is formed to minimize the distance from the presented stimuli, while the attribute-frequency model [@Neumann1974] states that prototype consist of attributes that were presented with highest frequency. Further studies showed higher experimental support for prototype-distance model for geometric patterns [@Franks1971] or faces [@Solso1979,@Solso1993,@Sampaio2018]. In the context of visual scenes, similar findings were not studied as it was difficult how to measure, what is a visual prototype.

Recent years brought possible solution to problem of visual prototype similarity of scenes. @Oliva2001 created a GIST descriptor that characterizes the spatial layout of the scene and it can be abstracted to natural properties of the scenes [@Torralba2003], which human observers can naturally use [@greene2009, @greene2010]. For this descriptor, we could compute pairwise distances, which would allows us to determine, which allows us to create an similarity space. With advances in training the convolutional neural networks [CNN, @Lecun2015,@goodfellow2016], they CNNs were widely used for various task in vision science in areas like scene categorization [@Krizhevsky2012], object recognition [@Kheradpisheh2016] or memorability of the image [@khosla2015]. One particular use of CNNs is to use the pretrained network, present it with a stimuli and use activation in one of the layers as a descriptor of the stimuli. The descriptors extracted in this way are knows as *deep features*. Again, we can compute distance between the deep features using L2 norm. @lukavsky2017 used this approach in a memory study to explore the properties of the similarity space defined by deep features.

@@

### Participants
Twelve subjects (mean age = 24.83, SD = 3.49, 6 males) participated in this experiment. All of them had normal or corrected-to-normal vision and none of that had participated in this type of experiment before. Subjects were students of Faculty of Arts and participated as a part of a course requirements. All participants gave an informed consent before the experiment. sample size was based on prior power analysis done in software G*Power 3.1.92 [@Faul2009]. We were aiming to detect large effect using one-sample t-tests while having $\alpha$=.05 and power $1-\beta$ = .8., which resulted in the above mentioned twelve participants.

@@

(ref::odd-scheme) Schema of scene selection. Green dot denotes selected scene, seven blue dots denotes distractors, and red dot denotes target scene. The representation is visualized in 2D for simplicity.

@@

Experiment was programmed in software Psychopy1.90.2 [@Peirce2019]. There were 64 trials in total. In each trial, nine scenes were shown in a rectangular grid (size of each scene was 256$\times$256 px). The scenes were selected as described in Design section. The position of all scenes in the grid was randomized. The task was to detect the scene that was most different from the rest. No additional instructions, which criteria should be taken into account for oddity were not given. Participants responded by mouse selection, the time for the response was unlimited. The experiment lasted approximately 20 minutes. We created six versions of the protocols and each protocols was supplied to two participants.

@@

### Accuracy for detecting different scene
The accuracy of detecting different scene significantly increased with increased distance ($\chi^2(1) = 21.7, p < .001$). As visualized in Figure \@ref(fig:odd-results) The accuracy for second quintile was not significant from the chance (chance level: 1/9, mean = 10%, SD = 9 %, _t_(11) = -0.316, _p_ = .758, _d_ = 0.09), while they were significant for third quintile (mean = 21%, SD = 9 %, _t_(11) = 3.9, _p_ = .003, _d_ = 1.12) and fourth quintile (mean = 26%, SD = 12 %, _t_(11) = 4.35, _p_ = .001, _d_ = 1.26). When the task was reparametrized using ratio of distances, we achieved same results ($\chi^2(1) = 30.4, p < .001$) as visualized in right part of Figure \@ref(fig:odd-results). 
In Figure \@ref(fig:odd-results-category), we can see accuracy for each category. The exploration of the image space for top and bottom categories did not reveal any obvious pattern. In order to do that, we reduced the data using multidimensional scaling. The image space for top category (field) and bottom category (closet) is visualized in Figure \@ref(fig:odd-results-MDS).
As each pair of participants received same set of images, we tested, whether the participants make the same pattern when selecting odd scene. We estimated the agreement in selection using Cohen's $\kappa$, Average agreement was .24 (SD = .06), which is considered as weak agreement [@fleiss2003]. Participants therefore selected different scenes as the most different ones.

(ref::odd-results) Average accuracy when detecting odd scene. Left plot shows accuracy for each quintile. Dot shows average accuracy, while vertical lines shows bootstrapped standard error of the mean. The distribution of the data is shown by violin plot. Right part shows fitted generalized linear mixed model when the distance was expressed using ratio of distances.

```{r odd-results, fig.cap = "(ref:odd-results)", fig.align = 'center'}
knitr::include_graphics(here("plots","Fig_odd_results.png"))

@@

(ref::odd-results-category) Average accuracy per each category. No clear trend for natural and man-made scene was observed. Vertical lines denote bootstrapped standard error of the mean.

@@

(ref::odd-results-MDS) Image space in MDS coordinates. Left shows category with lowest average accuracy, right shows category with highest average accuracy.

@@

Each of the descriptors was used as follows. For each 9 scenes, we computed pairwise L2 distances between corresponding descriptors resulting into distance matrix. From this structure, we evaluated the distances for each scene from the rest using hierarchical clustering as suggested by [@Torgo2007]. This method ranked the scenes in amount, how much they are distant from the rest. This selection was then compared with human selection of odd scene and with selection of distant scene by image space metric.

@@

(ref::odd-metrics) Prediction of odd scene using gist, HOG, SIFT, and RGB metrics. Left plots shows prediction of odd scene as defined using image space structure. Right part shows prediction of human performance.

@@

Taken together, even with the small sample size, we shown that participants are able to understand the distances defined by the image space and thus this experiment serves as validation experiment for further use of image spaces. Similar approach could be used for other perceptual spaces, defined either by CNN e.g. [@Groen2018, @lukavsky2017] or by multidimensional scaling [@Hout2014a].

@@

We tested the memory performance in an paradigm inspired by Deese-Roedrigez-McDermott paradigm [DRM, @Deese1959, @Roediger1995]. We presented the scenes and participants task was to memorize then. The similarity defined by image space serves as valid tool how to quantify similarity of stimuli and describe the presented stimuli using the coordinates in image space, where the center of presented stimuli should lie. we queried the subjects with either seen or unseen scenes, while we varied, whether the scenes were closer to the center of the presented scenes or further away. If the visual stimuli are internally represented in a structure showing resemblance to the image space, there should be more false alarms for the distractors closer to the center than for distractors further away from the center, as the center represent the visual prototype [@Posner1969].

@@

We selected the stimuli for the experiment from the database FIGRIM [@bylinskii2015], which contains 21 categories, each with more than 300 scenes of size 700$\times$700 px. Photographs of scenes in FIGRIM database come from larger SUN database [@xiao2010], which contain many artifacts (e.g. some scenes were not photographs of real scenes, but they were built from lego bricks, contain camera recording markers, etc.). Before the selection of the stimuli, we manually removed are scenes that had some artifacts in them. After the selection, 8665 photographs remained. For those scenes, we again evaluated each of them by pre-trained CNN and created an image space from the activation of fc7 layer.

@@

(ref::drm-scheme) Schema of experimental design. The black x denotes category center and dotted circle group of closer scenes. For randomly selected targets from the closer group (green dots), we compute the center (green x) and select close-distractors (red dots). From the farther scenes, we select far-distractors (blue dots).

@@

(ref::drm-procedure) Experimental procedure for one training and testing blog.

@@

We also analyzed the data using Signal detection theory approach [@green1966,@MacmillanNA2004]. Use of confidence ratings allowed us to construct empirical ROC curves with non-unit slopes for discriminating between targets and close-distractors and targets and far-distractors. We expressed the performance using $d_a$ measure as an alternative to traditional $d'$ used in a case of non-unit ROC slopes, which is recommend to be use in memory studies [@Loiotile2015]. Using this approach, the bias is not computed traditionally, but instead the the perceptual distance between scene types is computed [@MacmillanNA2004, @Loiotile2015].

@@

One sample t-tests revealed that recognition was above chance level in all three conditions (_p_s < .001, Cohen's _d_s > 1.52). The difference between two distractor types was significant ($\chi^2(1) = 56.1, _p_ < .001$)

(ref::drm-results-main) Accuracy results for each type of scene. The shape of data distribution is visualized using violin plot. The vertical line shows bootstrapped standard error of the mean, horizontal line denotes chance level. The false alarm rate for distractors could be computed by subtracting the accuracy from 100%.

@@

(ref::drm-results-ROC) ROC curves for discrimination between Target and close-distractors (left) and target and far-distractors (right). Black line denotes chance level. Red line shows average ROC curve, while the grey lines shows ROC curves for individual participants.

@@

(ref::drm-results-SDT-model) SDT model for three scene types. The perceptual distance between distractors was small (pseudo $d_a$ = 0.56).

@@

The exploration of recognition accuracy per each scene category revealed that easiest were categories bridge and bathroom, while most difficult were badlands and cockpit. Full results are displayed in Figure \@ref(fig:drm-results-category). On the level individual scene, we explored, whether the memory accuracy could be explained by memorability of the photograph [@Isola2001]. The memorability is good candidate to predict the task, because it has been shown that memorable images shows good perceptual organisation [@Goetschalckx2018] and are quickly categorisable [@Broers2017]. Therefore, we evaluated the photographs using trained CNN LaMem, which can predict memorability similarly to the behavioral scores [@khosla2015]. We then computed correlation of memory accuracy with this predicted memorability. The results showed that the correlation was small and significant for targets (_r_ = .10, _p_ = .003)) and for close-distractors(_r_ = .15, _p_ = .007)), while not significant for far-distractors (_r_ = .05, _p_ = .335). Therefore, the effect of memorability is only small.

(ref::drm-results-category) Moving average for reaction time (size of window was 200 consescutive trials) with respect to id of the trial.

@@

We were also interested, whether the recognition of the presented scenes propagated to the time needed to answer. We excluded the responses longer than 5s. Participants tended to make faster responses during the time course of the experiment as visualized in Figure \@ref(fig:drm-results-rt_window). As visualized in Figure \@ref(fig:drm-results-rt), the fastest responses were for far-distractors (mean = 1.09 s, SD = 0.29), followed by target scenes (mean = 1.19 s, SD = 0.21), while slowest responses were for close-distractors (mean = 1.26 s, SD = 0.21). The differences were significant ($\chi^2(1) = 27.8, _p_ < .001$). Pairwise post-hoc analysis with Tukey's correction revealed that the difference in reaction times between targets and close-distractors were not significant (_p_ = .256), while the remaining pairwise differences were significant (p < .001). Moreover, the difference in reaction times between close- and far-distractors was large (Cohen's d = 1.57).

(ref::drm-results-rt-window) Moving average for reaction time (size of window was 200 consescutive trials) with respect to id of the trial.

@@

(ref::drm-results-rt) Reaction times for each scene type. The shape of data distribution is visualized using violin plot. The vertical line shows bootstrapped standard error of the mean.

@@

(ref::drm-results-window-acc) Moving average (N = 200) for recognition accuracy. Accuracy for the distractors was stable, while the accuracy declined approximately after 1000 trial.

>>>>

author: 
  - name          : "Filip Děchtěrenko"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Hybernská 8, 11000, Prague 1"
    email         : "filip.dechterenko@gmail.com"
  - name          : "Jiří Lukavský"
    affiliation   : "2"
  - name          : "Jiří Štipl"
    affiliation   : "1"

@@

authornote: |
  Filip Děchtěrenko, Department of Psychology, Faculty of Arts, Charles University, Celetná 20, 110 00, Prague, Czech Republic. Institute of Psychology, Czech Academy of Sciences, Hybernská 8, 110 00, Prague, Czech Republic. Jiří Lukavský, Institute of Psychology, Czech Academy of Sciences, Hybernská 8, 110 00, Prague, Czech Republic. Jiří Štipl, Department of Psychology, Faculty of Arts, Charles University, Celetná 20, 110 00, Prague, Czech Republic.

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

World around us is full of information to process and store. When studying visual perception, it has been observed that visual memory has massive capacity for complex stimuli [@standing1970; @standing1973]. Recent studies showed that we are able to store not only shallow representation of the objects, but they are able to store many details [@brady2008] and similar findings were observed for scenes [@konkle2010b]. 
Although the capacity is large, it is an open question, how is it organised. The high recollection accuracy get lower when using yes/no task instead of 2AFC [@Cunningham2015]. Moreover, recent results show that the memory is much more limited in detail than it seems [@Draschkow2019].

How is the long-term memory organized? One possible explanation would be to create a coding scheme, using which the memories will be categorized [@brady2011], similarly as in working memory [@brady2009]. This coding scheme is created during our lifetime with the formation of semantic categories [@Kemp2009]. In the case of objects, the possible candidate is a hierarchical structure from simple features through larger meaningful groups to whole objects [@brady2011; @Ullman2007]. In the case of scenes is the situation more complicated as not only the objects in the scene should be represented, but also their spatial arrangement and the gist of the scene [@Oliva2005]. The organisation of whole scenes in the memory remains an open question.

One possible approach to this problem could be through formation of categories. The semantic categories in our mind are formed to match the categories of outside world and to be economic with terms of used properties to describe the structure [@rosch1978]. The category is formed around a prototype - a typical exemplar, which shows the maximal resemblance with rest of the members in the category, while having a low resemblance with other categories [@rosch1975b]. The prototypes are created not only for semantic categories, but also for visual stimuli. There are two main theories around visual prototype formation. The prototype-distance model [@Posner1970] states that the prototype is formed to minimize the distance from the presented stimuli, while the attribute-frequency model [@Neumann1974] states that prototype consist of attributes that were presented with highest frequency. Further studies showed higher experimental support for prototype-distance model for geometric patterns [@Franks1971] or faces [@solso1979; @Solso1993; @Sampaio2018]. In the context of visual scenes, similar findings were not studied as it was difficult how to measure, what is a visual prototype.

Recent years brought possible solution to problem of visual prototype similarity of scenes. @Oliva2001 created a GIST descriptor that characterizes the spatial layout of the scene and it can be abstracted to natural properties of the scenes [@Torralba2003], which human observers can naturally use [@greene2009; @greene2010]. For this descriptor, we could compute pairwise distances, which would allows us to create a similarity space. With advances in training the convolutional neural networks [CNN, @Lecun2015; @goodfellow2016], the CNNs were adopted for various task in vision science like scene categorization [@Krizhevsky2012], object recognition [@Kheradpisheh2016] or memorability of the image [@khosla2015]. One particular use of CNNs is evaluate set of stimuli using pretrained network, take output of one of the inner layers and use it as a descriptor of the stimuli. The descriptors extracted in this way are known as *deep features*. Again, we can compute distance between the deep features using L2 norm. @lukavsky2017 used this approach in a memory study to explore the properties of the similarity space defined by deep features.

@@

### Participants
Twelve subjects (mean age = 24.83, SD = 3.49, 6 males) participated in this experiment. All of them had normal or corrected-to-normal vision and none of that had participated in this type of experiment before. Subjects were students of Faculty of Arts and participated as a part of a course requirements. All participants gave an informed consent before the experiment. sample size was based on prior power analysis done in software G*Power 3.1.92 [@Faul2009]. We were aiming to detect large effect using one-sample t-tests while having $\alpha$=.05 and power $1-\beta$ = .8, which resulted in the above-mentioned twelve participants.

@@

(ref:odd-scheme) Schema of scene selection. Green dot denotes selected scene, seven blue dots denotes distractors, and red dot denotes target scene. The representation is visualized in 2D for simplicity.

@@

Experiment was programmed in software Psychopy1.90.2 [@Peirce2019]. There were 64 trials in total. In each trial, nine scenes were shown in a rectangular grid (size of each scene was 256 $\times$ 256 px). The scenes were selected as described in Design section. The position of all scenes in the grid was randomized. The task was to detect the scene that was most different from the rest. No additional instructions, which criteria should be taken into account for oddity were not given. Participants responded by mouse selection, the time for the response was unlimited. The experiment lasted approximately 20 minutes. We created six versions of the protocols and each protocols was supplied to two participants.

@@

### Accuracy for detecting different scene
The accuracy of detecting different scene significantly increased with increased distance ($\chi^2(1) = 21.7, p < .001$). As visualized in Figure \@ref(fig:odd-results), the accuracy for second quintile was not significant from the chance (chance level: 1/9, mean = 10%, SD = 9 %, _t_(11) = -0.316, _p_ = .758, _d_ = 0.09), while they were significant for third quintile (mean = 21%, SD = 9 %, _t_(11) = 3.9, _p_ = .003, _d_ = 1.12) and fourth quintile (mean = 26%, SD = 12 %, _t_(11) = 4.35, _p_ = .001, _d_ = 1.26). When the task was reparametrized using ratio of distances, we achieved same results ($\chi^2(1) = 30.4, p < .001$) as visualized in right part of Figure \@ref(fig:odd-results). 
In Figure \@ref(fig:odd-results-category), we can see accuracy for each category. The exploration of the image space for top and bottom categories did not reveal any obvious pattern. In order to do that, we reduced the data using multidimensional scaling. The image space for top category (field) and bottom category (closet) is visualized in Figure \@ref(fig:odd-results-MDS).
As each pair of participants received same set of images, we tested, whether the participants make the same pattern when selecting odd scene. We estimated the agreement in selection using Cohen's $\kappa$, Average agreement was .24 (SD = .06), which is considered as weak agreement [@fleiss2003]. Participants therefore selected different scenes as the most different ones.

(ref:odd-results) Average accuracy when detecting odd scene. Left plot shows accuracy for each quintile. Dot shows average accuracy, while vertical lines shows bootstrapped standard error of the mean. The distribution of the data is shown by violin plot. Right part shows fitted generalized linear mixed model when the distance was expressed using ratio of distances.

```{r odd-results, fig.cap = "(ref:odd-results)", fig.align = 'center'}

@@

(ref:odd-results-category) Average accuracy per each category. No clear trend for natural and man-made scene was observed. Vertical lines denote bootstrapped standard error of the mean.

@@

(ref:odd-results-MDS) Image space in MDS coordinates. Left shows category with lowest average accuracy, right shows category with highest average accuracy.

@@

Each of the descriptors was used as follows. For each 9 scenes, we computed pairwise L2 distances between corresponding descriptors resulting into distance matrix. From this structure, we evaluated the distances for each scene from the rest using hierarchical clustering as suggested by [@torgo2007]. This method ranked the scenes in amount, how much they are distant from the rest. This selection was then compared with human selection of odd scene and with selection of distant scene by image space metric.

@@

(ref:odd-metrics) Prediction of odd scene using gist, HOG, SIFT, and RGB metrics. Left plots shows prediction of odd scene as defined using image space structure. Right part shows prediction of human performance.

@@

Taken together, even with the small sample size, we shown that participants are able to understand the distances defined by the image space and thus this experiment serves as validation experiment for further use of image spaces. Similar approach could be used for other perceptual spaces, defined either by CNN e.g. [@Groen2018; @lukavsky2017] or by multidimensional scaling [@Hout2014a].

@@

We tested the memory performance in an paradigm inspired by Deese-Roedrigez-McDermott paradigm [DRM, @Deese1959, @Roediger1995]. We presented the scenes and participants task was to memorize then. The similarity defined by image space serves as valid tool how to quantify similarity of stimuli and describe the presented stimuli using the coordinates in image space, where the center of presented stimuli should lie. we queried the subjects with either seen or unseen scenes, while we varied, whether the scenes were closer to the center of the presented scenes or further away. If the visual stimuli are internally represented in a structure showing resemblance to the image space, there should be more false alarms for the distractors closer to the center than for distractors further away from the center, as the center represent the visual prototype [@Posner1970].

@@

We selected the stimuli for the experiment from the database FIGRIM [@bylinskii2015], which contains 21 categories, each with more than 300 scenes of size 700 $\times$ 700 px. Photographs of scenes in FIGRIM database come from larger SUN database [@xiao2010], which contain many artifacts (e.g. some scenes were not photographs of real scenes, but they were built from lego bricks, contain camera recording markers, etc.). Before the selection of the stimuli, we manually removed are scenes that had some artifacts in them. After the selection, 8665 photographs remained. For those scenes, we again evaluated each of them by pre-trained CNN and created an image space from the activation of fc7 layer.

@@

(ref:drm-scheme) Schema of experimental design. The black x denotes category center and dotted circle group of closer scenes. For randomly selected targets from the closer group (green dots), we compute the center (green x) and select close-distractors (red dots). From the farther scenes, we select far-distractors (blue dots).

@@

(ref:drm-procedure) Experimental procedure for one training and testing blog.

@@

We also analyzed the data using Signal detection theory approach [@green1966; @MacmillanNA2004]. Use of confidence ratings allowed us to construct empirical ROC curves with non-unit slopes for discriminating between targets and close-distractors and targets and far-distractors. We expressed the performance using $d_a$ measure as an alternative to traditional $d'$ used in a case of non-unit ROC slopes, which is recommend to be use in memory studies [@Loiotile2015]. Using this approach, the bias is not computed traditionally, but instead the the perceptual distance between scene types is computed [@MacmillanNA2004; @Loiotile2015].

@@

One sample t-tests revealed that recognition was above chance level in all three conditions (_p_s < .001, Cohen's *d*s > 1.52). The difference between two distractor types was significant ($\chi^2(1) = 56.1, p < .001$)

(ref:drm-results-main) Accuracy results for each type of scene. The shape of data distribution is visualized using violin plot. The vertical line shows bootstrapped standard error of the mean, horizontal line denotes chance level. The false alarm rate for distractors could be computed by subtracting the accuracy from 100%.

@@

(ref:drm-results-ROC) ROC curves for discrimination between Target and close-distractors (left) and target and far-distractors (right). Black line denotes chance level. Red line shows average ROC curve, while the grey lines shows ROC curves for individual participants.

@@

(ref:drm-results-SDT-model) SDT model for three scene types. The perceptual distance between distractors was small (pseudo $d_a$ = 0.56).

@@

The exploration of recognition accuracy per each scene category revealed that easiest were categories bridge and bathroom, while most difficult were badlands and cockpit. Full results are displayed in Figure \@ref(fig:drm-results-category). On the level individual scene, we explored, whether the memory accuracy could be explained by memorability of the photograph [@Isola2011]. The memorability is good candidate to predict the task, because it has been shown that memorable images shows good perceptual organisation [@Goetschalckx2018] and are quickly categorisable [@Broers2017]. Therefore, we evaluated the photographs using trained CNN LaMem, which can predict memorability similarly to the behavioral scores [@khosla2015]. We then computed correlation of memory accuracy with this predicted memorability. The results showed that the correlation was small and significant for targets (_r_ = .10, _p_ = .003)) and for close-distractors(_r_ = .15, _p_ = .007)), while not significant for far-distractors (_r_ = .05, _p_ = .335). Therefore, the effect of memorability is only small.

(ref:drm-results-category) Moving average for reaction time (size of window was 200 consescutive trials) with respect to id of the trial.

@@

We were also interested, whether the recognition of the presented scenes propagated to the time needed to answer. We excluded the responses longer than 5s. Participants tended to make faster responses during the time course of the experiment as visualized in Figure \@ref(fig:drm-results-rt_window). As visualized in Figure \@ref(fig:drm-results-rt), the fastest responses were for far-distractors (mean = 1.09 s, SD = 0.29), followed by target scenes (mean = 1.19 s, SD = 0.21), while slowest responses were for close-distractors (mean = 1.26 s, SD = 0.21). The differences were significant ($\chi^2(1) = 27.8, p < .001$). Pairwise post-hoc analysis with Tukey's correction revealed that the difference in reaction times between targets and close-distractors were not significant (_p_ = .256), while the remaining pairwise differences were significant (p < .001). Moreover, the difference in reaction times between close- and far-distractors was large (Cohen's d = 1.57).

(ref:drm-results-rt-window) Moving average for reaction time (size of window was 200 consescutive trials) with respect to id of the trial.

@@

(ref:drm-results-rt) Reaction times for each scene type. The shape of data distribution is visualized using violin plot. The vertical line shows bootstrapped standard error of the mean.

@@

(ref:drm-results-window-acc) Moving average (N = 200) for recognition accuracy. Accuracy for the distractors was stable, while the accuracy declined approximately after 1000 trial.
